### 2014-08-21 tech.tex
now

\section{Basics}~\label{sec:basic}

\subsection{Trace Terminology}
Our analysis starts with a trace $\tau$,  a sequence of events, $e_0, e_1, \dots, e_n$.  
There are three types of events in general.
\begin{itemize}
\item the shared access, which includes the read and write of the shared fields, e.g., $o.f$=$x$ and $x$=$o.f$.
\item the local access, which includes only the access of the local variable, e.g., $x=y+z$ or $v.f=x$ (where $v$ is a thread-local object).
\item the branch event, which evaluates the branch condition to true/false, e.g., $x>3$.
\item the synchronization event, which includes start/join, wait/notify, lock/unlock events, e.g., $lock(o)$
\end{itemize}

Each event $e_i \in \Sigma$ is a tuple, $<t, id, a, v, ins>$, where $t\in \mathcal{T}$ denotes the thread generating the event, $id\in \mathcal{ID}$ denotes the unique integer assigned to the event (event id), $a\mathcal{A}$ denotes the address of the object or field (if any) accessed in the event,  $v\in \mathcal{V}$ denotes the value of the definition (if any) in the event, and $ins\in \mathcal{INS}$ denotes the three-address instruction generating the event.  Specifically, the address of the object $o$ is denoted as $id(o)\in \mathcal{ID}$, which is a string value representing $o$ uniquely, the address of the static field $f$ is  $id(f)$ and the address of the instance object field $o.f$ is  $id(o)\_id(f)$. Besides, as the event is derived by instrumenting the three-address code and monitoring the instrumented execution. Therefore, each event can involve at most three operands.

The trace supports its standard operations as follows.
\begin{itemize}
\item projection, e.g., $\tau|t$ returns~\footnote{This is the abbreviation for the complete form $\tau|Thread=t$ } the events from the thread $t$,  $\tau|a$ returns the event involving the address $a$.  
\item concatenation. $\tau'=\tau e$ represents the new trace by appending the event $e$ to $\tau$.
\item length $|\tau|$. 
\item selecting an element. $\tau[0]$ and $\tau[|\tau|-1]$ represents the first and last event in $\tau$.
\end{itemize}

%The modeling of synchronization event is standard and explained in existing work, so we focus on the rest two types of events in this paper.

In addition, we maintain auxiliary information as follows.
\begin{itemize}
\item $AT: \mathcal{A} \times \mathcal{T} \rightarrow \gamma$ is a function that returns a trace $\tau \in \gamma$ that contains only the accesses of the address $a \in \mathcal{A}$ by the thread $t\in \mathcal{T}$. Each trace $\tau$ in $\gamma$ is defined over the alphabet of events $\Sigma$, specifically, the trace is an empty trace $\epsilon$ or defined in this way:  $\forall 0\leq i\leq |\tau|,   \tau[i]\in \Sigma, and, \forall i\neq j, \tau[i]\neq \tau[j]$. 
\item $R: \mathcal{A} \rightarrow \gamma$ is a function that returns the read accesses of the address $a\in \mathcal{A}$.
\item $W: \mathcal{A} \rightarrow \gamma$ is a function that returns the write accesses of the address $a \in \mathcal{A}$.
\item $Sync: \mathcal{A} \rightarrow \gamma$ is a function that returns the synchronization events involving the address $a\in \mathcal{A}$. 
\end{itemize}

\subsection{Symbolic Trace}
To facilitate the symbolic analysis, we need to introduce symbols to represent the operands in each event. Symbols allow us to overcome the limitation of concrete dependences and allow us to explore more dependences symbolically.

%shared access only, local access only

{\bf Local Variables\  }
Like other symbolic analysis~\cite{jeff,chao}, the symbolic trace should be in the SSA form, i.e., each variable is defined exactly once. This is because the constraint solver employed by the analysis requires each variable to hold only one value. Besides, we need to make sure each use still reads from the same definition thread-locally.

The  simple procedure shows the construction of the symbolic trace for local variables defined or used. The symbols are constructed by combining the static instruction and the runtime event id. Lines 6-10 handles the local variable definition.  We build a symbolic variable $s$ for it by combining the variable name and the event id. The uniqueness of the event id guarantees that each symbolic variable is defined exactly once. In addition, we replace the variable to the symbolic variable in the instruction and record the replacement in $table$. Lines 3-5 updates the local variable used in each event so that it is replaced with the symbolic variable for the corresponding definition. Here, the corresponding definition and the use share the same variable name, therefore, we can easily find out the symbolic variable through looking up the $table$.

The SSA form of the trace is different from the SSA form of the instruction as the SSA instruction can only distinguish definitions at different program points but cannot distinguish the definitions at different execution points that share the same program point.

\begin{algorithmic}[3]
\For {$e: \tau$}
 \State $ins\gets e.ins$ 
 \For {$ins.use:ins.uses$}
 \State $ins.use \gets table(ins.use)$
 \EndFor
  \If {$ins.def\neq null$}
    \State $ s\gets ins.def^{e.id}$
	\State $ table [ins.def \rightarrow s]$
	\State $ins.def\gets s$
 \EndIf
\EndFor
\end{algorithmic}

%TODO update the intro+moti, make sure the same style
\begin{figure}
\centering
\begin{tabular}{ll}
\multicolumn{2}{c}{{\tt {\bf x} = 0; {\bf y} = 0;}} \\
\hline
\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
\hline
{\tt 1: s=0; } & \\
{\tt 2: for(i=1;i<3;i++)} & \\
{\tt 3: \ \ \ s+=i;} & \\
{\tt 4: {\bf y} = s;} & \\
{\tt 5: {\bf x} = 1;} & \\
{\tt 6: {\bf y} = 5;} & \\
& {\tt 7: if ({\bf y} > 2)} \\
& {\tt 8:~~print({\bf x}+1);} \\	
& {\color{Gray} {\tt 9: else}} \\
& {\color{Gray} {\tt 10:~~print({\bf x}+2);}}
\end{tabular}
\caption{Running Example (shared variables are in bold font). }
\label{fig:running2}
\end{figure}

{\bf Shared Accesses}   Besides, we introduce symbols to represent shared reads and shared writes which leave the inter-thread dependence between reads/writes undetermined. For each read (or write) of shared variable $x$, we introduce $R^{id}_x$ (or $W^{id}_x$) to denote it, where $id$ is the event id.
Consider the code in Figure~\ref{fig:running2}, which resembles the example in Figure~\ref{fig:running} except that it includes a for loop at lines 1-2.
The symbolic trace is produced in Figure~\ref{fig:t4running2}. For simplicity, we only show the symbolic variables, while omitting other information such as thread information.

%TODO distinguish T1 and T2 in the text. 
\begin{figure}
\centering
\begin{tabular}{l|l}
\hline
\multicolumn{1}{c}{$Trace$} & \multicolumn{1}{c}{$Symbolic\  Trace$} \\
\hline
{\tt 0: {\bf x}=0} &  {\tt 0: $W^0_x$=0}    \\
{\tt 1: {\bf y}=0} &   {\tt 1: $W^1_y$=0}   \\
{\tt 2: s=0} &  {\tt 2: $s^2$=0}   \\
{\tt 3: i=1} &     {\tt 3: $i^3$=1}   \\
{\tt 4: i<3} &    {\tt 4: $i^3$<3} \\
{\tt 5: s=s+i} & {\tt 5: $s^5$=$s^2$+$i^3$}   \\
{\tt 6: i=2} &       {\tt 6: $i^6$=2}  \\
{\tt 7: i<3} &      {\tt 7: $i^6$<3}  \\
{\tt 8: s=s+i} &  {\tt 8: $s^8$=$s^5$+$i^6$}  \\
{\tt 9: i=3} &     {\tt 9: $i^9$=3}  \\
{\tt 10: i<3} &    {\tt 10: $i^9$<3}  \\
{\tt 11: {\bf y} = s;} &  {\tt 11: $W^{11}_y$ = $s^8$;}  \\
{\tt 12: {\bf x} = 1;} &    {\tt 12: $W^{12}_x$ = 1;}   \\
{\tt 13: {\bf y} = 5;} &    {\tt 13: $W^{13}_y$ = 5;}  \\
{\tt 14: {\bf y} > 2}  &    {\tt 14: $R^{14}_y$ > 2} \\
{\tt 15: tmp={\bf x}+1;}  & {\tt 15: tmp=$R^{15}_x$+1;}   \\	
{\tt 16: print(tmp);} &  {\tt 16: print(tmp);}  \\
\end{tabular}
\caption{Trace}
\label{fig:t4running2}
\end{figure}

{\bf Method Calls\ } The key to supporting the method calls is to capture the value flow the actual argument to the formal argument, and the flow from the return statement to the LHS variable of the method call. To explicitly model the value flow, record two additional events for each method call.
Consider the example in Figure~\ref{fig:methcall},   we record the local access event $y1=y;$ for the argument value flow and record the  local access event $x=i2;$. Recording the additional events is achieved through instrumenting the call site and the callee method statically.

The above simple strategy however hides the complexity of the virtual method calls. At a call site of a virtual method, the static instrumentation cannot know precisely which method would be called. Therefore, we do not know what formal argument the actual argument flows to. Consider the example in Figure~\ref{fig:methcall}, suppose another implementation of the virtual method exists (in the comments).  We do not know how to instrument the code statically, $y1=y$ or $y2=y$.

To avoid the problem, we have to combine the runtime knowledge. 
Our strategy is as follows: rather than directly record direct value flow from actual argument to formal argument, we introduce an artificial variable during the static instrumentation. Then we  insert the instrumentation $record(ARG0=y;)$ at call site, and insert  $record(y1=ARG0)$ at the entry of the method $func$ declared in the first class, and insert $record(y2=ARG0);$ at the entry of the method $func$ declared in the second class. At runtime, depending on which $func$ method is invoked, we record either the event sequence $ARG0=y; y1=ARG0$ or the sequence $ARG0=y; y2=ARG0$, which precisely captures the value flow. We model the return value flow similarly.

Note that although different methods use the same names for the artificial variable,  they are translated to different variables after we get the SSA form of symbolic trace.

\begin{figure}
\centering
\begin{tabular}{ll}
{\tt  x=o.func(y) } &  \\ 
 {\tt  func(y1)\{ // class O1} &  \\
 {\tt  \ \      i=y1; } & \\
 {\tt  \ \      i2=2*i; } &  \\
 {\tt  \  \      return i2;} & \\ 
 {\tt           \}} & \\

  {\tt //  func(y2)\{ // class O2} &  \\
 {\tt  // \ \      j=y2; } & \\
 {\tt  // \ \      j2=3*j; } &  \\
 {\tt  // \  \      return j;} & \\ 
 {\tt  //         \}} & \\ 
\end{tabular}
\caption{Method Calls}
\label{fig:methcall}
\end{figure}

\subsection{Constraints}
The constraints

\section{Time Window}~\label{sec:basic}

\section{Heap Invariant}~\label{sec:basic}