### 2014-09-02 introduction.tex
intro

\section{Introduction}\label{Se:introduction}

In a multithreaded application, a data race occurs when two concurrent threads access the same memory location, such that (i) at least one of the accesses is write access, and (ii) no explicit mechanism is enforced to prevent the threads from simultaneous access to the location \cite{eraser}. Race detection is important not only because races often reveal bugs in the implementation, but also as the basis of other techniques and analyses like atomicity checking \cite{learning,avio,ordervio}, data-flow analysis \cite{dataflow} and record/replay \cite{leap,replay2012}.

In this paper, we focus on \emph{sound} race detection, whereby reported races are guaranteed to be real. This is the key requirement for adoption of the tool by developers \cite{XXX}. Ensuring soundness is a difficult challenge, which mandates dynamic forms of analysis. Indeed, extensive research has been carried out on dynamic race detection. The general goal has been to derive constraints from a given execution trace on event reordering, and check for the remaining reorderings whether they disclose data races. The key question then becomes about constraint extraction.

%Fasttrack PLDI paper claims happens-before analysis never reports false positives. Check.
\paragraph{Existing Approaches} 
Numerous techniques have been proposed to date for race detection, which either sacrifice soundness~\cite{ucg,XXX,YYY,ZZZ}  or have significant coverage limitations~\cite{XXX,YYY,ZZZ}. Recently \emph{predictive analysis} has emerged as a promising alternative, whereby a single trace is considered, and races are discovered by permuting the execution schedule governing the trace \cite{JEFF,YANNIS}.

Predictive analysis guarantees soundness, and has the potential for high coverage, though current predictive analyses suffer from two major limitations:
\begin{enumerate}
	\item All the permutations attempted by the analysis must preserve the flow dependencies exhibited by the original trace. As an example, if in the original trace thread $t_1$ reads a shared variable that was written by thread $t_2$, then the same must hold in the permuted trace.
	\item The analysis cannot step outside the boundaries of the input trace, e.g. by exploring branches that were not followed in that trace.
\end{enumerate}
Both of these constraints are a conservative means of ensuring soundness. In an empirical study we conducted, which we describe in Section \ref{sec:eval}, we found that these restrictions often lead to serious loss in coverage.

As an illustration, we refer to the trace in Figure \ref{fig:running}, where the statements in red denote an unexplored branch outside the trace. Existing predictive analyses are unable to detect the race between lines {\tt 2} and {\tt 5} because of the dependence between lines {\tt 3} and {\tt 4}, which is a barrier to the needed reorderings. A second race that is missed, as it involves the statements in red, is between lines {\tt 2} and {\tt 7}.

%Initial attempts to address this challenge focused on built-in synchronization primitives. These include locks as well as the wait/notify and start/join scheduling controls. Notable among these efforts is \emph{lockset} analysis, which considers only locks. Because the derived constraints are partial, permitting certain infeasible event reorderings, lockset analysis cannot guarantee soundness. 
%
%A different tradeoff is struck by the \emph{happens-before (HB)} approach. In this style of analysis, all synchronization primitives are accounted for, though reordering constraints are conservative. As an example, HB inhibits reordering of two synchronized blocks governed by the same lock. Recently there have been successful attempts to relax HB constraints. Among these are hybrid analysis, which permits both orderings of lock-synchronized blocks, and the \emph{Universal Causal Graph (UCG)}~\cite{ucg} representation, which also enables both orderings but only if these are consistent with wait/notify- and start/join-induced constraints.
%
%Unfortunately, even full consideration of synchronization constructs is insufficient. Branching decisions must also be respected to ensure the feasibility of a reported race. As an example, we refer to the trace in Figure \ref{fig:running}. (For now, ignore the statements in red.) This trace is free of any explicit synchronization statements, and so a race between the assignment to variable {\tt x} at line {\tt 2} and the use of {\tt x} at line {\tt 5} may appear possible. This race is, however, spurious, since the use of {\tt x} at line {\tt 5} is governed by positive evaluation of the test at line {\tt 4}, which is conditioned on the assignment at line {\tt 3}. Hence, line {\tt 2} necessarily preceds line {\tt 5}.

\begin{figure}
\centering
\begin{tabular}{ll}
\multicolumn{2}{c}{{\tt x = 0; y = 0; z = 3;}} \\
\multicolumn{2}{c}{{\color{red} {\tt z = 2;}}} \\
\hline
\multicolumn{1}{c}{$T_1$} & \multicolumn{1}{c}{$T_2$} \\
\hline
{\tt 1: y = 3;} & \\
{\tt 2: x = 1;} & \\
{\tt 3: y = 5;} & \\
& {\tt 4: if (y > z)} \\
& {\tt 5:~~print(1/x);} \\	
& {\color{red} {\tt 6: else}} \\
& {\color{red} {\tt 7:~~print(2/x);}}
\end{tabular}
\caption{Example illustrating ordering constraints beyond synchronization primitives}
\label{fig:running}
\end{figure}

%As an alternative, driven by the requirement for soundness, 
%Smagardakis et al. have recently proposed \emph{predictive analysis}. In this approach, both synchronization constraints and inter-thread dependencies are preserved, where inter-thread dependencies are respected by 
%only allowing reorderings that leave the dependence structure exhibited by the original trace in tact. This ensures that the values of shared memory locations remain the same, which secures the soundness argument. As a soundness-preserving relaxation, Huang et al. \cite{XXX} permit reorderings as long as control dependencies are respected.
%
%While being sound and able to find real races, existing predictive analyses are conservative in two respects:
%\begin{enumerate}
%	\item Constraints imposed by branch conditions are abided by preserving the exact values flowing into the condition as in the original trace.
%	\item A second limitation, implied by the first, is that traces that diverge from the original trace in their branching history are strictly out of scope. 
%\end{enumerate}
% Due to these restrictions, enforced as a conservative means of ensuring soundness, existing predictive analysis suffer from limited coverage.

 \paragraph{Our Approach} We describe a novel approach to predictive race detection, which we have implemented as the \tool\ analysis tool, that is able to relax both of these constraints. Compared to the state of the art \cite{JEFF}, \tool\ is able to detect x$1.5$ more races with only the first restriction relaxed, and x$2.5$ more races with both restrictions relaxed.

 The departure point of our approach is to consider the explicit values of shared memory locations instead of dependencies between trace events. The gain in coverage is twofold:
 \begin{enumerate}
 	\item Reorderings that violate the original dependence structure but preserve the branching history become possible. The values read by a branching statement may change so long as the branch condition evaluates to its original truth value.
 	\item In addition, in certain cases we can guarantee soundness while directing execution toward branches that diverge from the original trace. This mandates that the effects of an unexplored branch can be modeled precisely, which holds frequently in practice because the original and new branchees often access the same set of shared memory locations.
 \end{enumerate}

 Returning to the example in Figure \ref{fig:running}, we demonstrate how value-based reasoning improves coverage. First, value-based reasoning reveals that trace 
 [{\tt 1}, {\tt 4}, {\tt 5}, {\tt 2}, {\tt 3}] is feasible, since the first assignment, {\tt y = 3}, satisfies the condition {\tt y > z}. Hence the race is discovered. A second race, between lines {\tt 2} and {\tt 7}, is detected by negating the condition. This is possible because the effects of the {\tt else} branch can be modeled precisely.
 Negation leads to an execution starting at line {\tt 4} (where {\tt y $\equiv$ 0}). The race between lines {\tt 2} and {\tt 7} then becomes visible.

 Concretely, we achieve value-level (rather than dependence-level) granularity computationally via a unique encoding of the input execution trace as a constraint system. The constraints are then processed by a satisfiability checker. This gives us the flexibility to explore nontrivial constraints, such as 
 forcing a context switch after the first assignment to {\tt y} or negating the condition in Figure \ref{fig:running}. Importantly, these manipulations are beyond what dependence-based reasoning can achieve, since this view of the trace is too conservative.

\paragraph{Contributions} This paper makes the following principal contributions: