### 2014-09-06 abstract.tex
minor

\begin{abstract}
Predictive analysis has recently emerged as a promising technique for 
sound detection of multithreading bugs (i.e., without false alarms). 
The idea is to start from an instrumented execution trace, and transform 
it based on the concrete information it provides, such as concrete values 
and memory accesses, to expose bugs, while preserving feasibility. For
example, the analysis may reorder trace events to simulate different thread schedules, 
as long as value flow dependencies are preserved.

Predictive analysis has been successfully applied to data-race detection, 
although the state-of-the-art techniques are overly restrictive in the 
transformations they allow, sacrificing coverage in return for a conservative soundness guarantee.

In this paper, we present \sysname, a predictive race detector that has two
important relaxations: value and path relaxations without affecting the
soundness guarantee. The former allows 
variables to have different values and the latter allows exploring 
unexecuted paths that are neighbors to the traced path.
Our results show that \sysname\ is able to detect $2.5X$ as many data races on a 
representative set of large-scale benchmarks as compared to the state-of-the-art.
%in predictive race detection.
\end{abstract}